
\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{bm}
\usepackage{physics}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage[numbers]{natbib}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{xcolor}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  citecolor=blue!60!black,
  urlcolor=blue!60!black
}

\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}

% --- Notation ---
\newcommand{\Alive}{\textsf{ALIVE}}
\newcommand{\Jam}{\textsf{JAM}}
\newcommand{\Mem}{\textsf{MEM}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\T}{\mathbb{T}}
\newcommand{\angleset}{\{0,90,180,270\}^\circ}
\newcommand{\Stress}{\mathrm{StressIndex}}
\newcommand{\Resist}{\mathrm{Resistance}}
\newcommand{\Wind}{\mathrm{Winding}}
\newcommand{\GLMS}{\mathrm{GLMS}}
\newcommand{\sgn}{\operatorname{sgn}}

\title{\textbf{The Logic Evaluation Engine (LEE)}\\
\large Toroidal Phase Geometry, Phase Tensors, and Stress-Aware Inference \emph{independently developed, later recognized as convergent with neuroscience}}
\author{Alexander Patterson}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
The Logic Evaluation Engine (LEE) is a symbolic inference system whose discrete phase dynamics (\Alive,\ \Jam,\ \Mem) produce closed cyclic flows on a low-dimensional manifold.
The toroidal structure observed in LEE predates---and was \emph{not} inspired by---neuroscience reports of low-dimensional toroidal activity \citep{Nature_2011,Nature_2024}; the similarity was noted ex post.
We present a constructive account of why a two-torus $\T^2$ (and higher tori) arise from material-implication rotations and contradiction archival, define a Global Logical Memory Space (\GLMS), and introduce operational metrics (Winding, Resistance, and a StressIndex) that quantify manifold distortion and system ``health.''
We give falsifiable predictions, an executable instrumentation plan, and show how LEE can model configurations \emph{beyond presently known biological constraints}, offering hypotheses testable in neuroscience and medicine.
\end{abstract}

\section{Introduction: independence, then convergence}
LEE was built from first principles in logic and phase geometry. Only later did external readers point out its resemblance to toroidal neural manifolds reported in \emph{Nature} (2011; 2024).
Our claim is therefore \emph{constructive}: toroidal structure is a necessity of LEE's operators, not an import from biology. That such manifolds also appear in cortical data suggests convergent constraints across information-processing substrates.

\paragraph{Contributions.}
(i) A minimal discrete-time model that yields $\T^2$ from logical first principles; (ii) a tensorial memory field (\GLMS) that explains curvature and loop persistence; (iii) stress-aware metrics with precise measurement recipes and provenance hooks; (iv) a protocol to compare LEE manifolds against neural-manifold datasets without tuning to biology.

\section{Formal model}
\begin{definition}[Phase alphabet and angles]
Let $\mathcal{S}=\{\Alive,\Jam,\Mem\}$ with angle map $\varphi:\mathcal{S}\!\to\!\{0,\tfrac{\pi}{2},\pi,\tfrac{3\pi}{2}\}$.
A run produces a sequence $(s_t)_{t=0}^T$, $s_t\in\mathcal{S}$.
The rotation at step $t$ is $\Delta\theta_t=(\varphi(s_t)-\varphi(s_{t-1}))\bmod 2\pi$, constrained to $\{0,\tfrac{\pi}{2},\pi,\tfrac{3\pi}{2}\}$.
\end{definition}

\begin{definition}[Global Logical Memory Space (\GLMS)]
Each transition emits a local tensor $\bm{g}_t\in\R^{k\times k}$ encoding contradiction strain (e.g., unmatched premises, unresolved constraints).
The \emph{phase tensor} of a run is $G=\sum_{t=1}^T \bm{g}_t$; archival in \Mem\ accumulates $G$, replay exposes $G$ for audit.
\end{definition}

\begin{proposition}[Two independent cycles $\Rightarrow$ $\T^2$]\label{prop:torus}
Suppose (i) evaluation rotates on $\theta\in S^1$ via $\Alive\!\rightarrow\!\Jam\!\rightarrow\!\Mem\!\rightarrow\!\Alive$; (ii) memory accrual rotates on an independent angle $\psi\in S^1$ via archival/replay.
Then the attractor manifold generically factorizes as $\T^2=S^1_\theta\times S^1_\psi$ up to gluing along identification sets where contradictions are resolved.
\end{proposition}

\begin{remark}[Beyond $\T^2$]
Adapters or side-cycles add independent $S^1$ factors (yielding $\T^k$); nontrivial gluing changes genus—LEE can thus exceed cortical constraints while preserving discrete stability.
\end{remark}

\section{Manifold distortion and operational metrics}
Let $\mathcal{M}_\mathrm{ideal}$ denote the ideal torus with major/minor radii $(R,r)$.
For an embedded cloud $\{\mathbf{x}_i(t)\}$ obtained from LEE state vectors (or phase traces embedded via spectral/PH methods), define:

\paragraph{Radial distortion.}
\begin{equation}
D_r(t)=\frac{1}{N}\sum_{i=1}^N \left(\,\big|\|\Pi_{\mathrm{maj}}\mathbf{x}_i(t)\|-R\big|+\big|\|\Pi_{\mathrm{min}}\mathbf{x}_i(t)\|-r\big|\,\right).
\end{equation}

\paragraph{Curvature concentration.}
Let $\kappa(\cdot)$ be an estimator of geodesic curvature along trajectories on $\mathcal{M}$. Define $D_\kappa(t)=\operatorname{IQR}\{\kappa(\gamma_i(t))\}$ over run segments $\gamma_i$.

\paragraph{Winding, Resistance, StressIndex.}
\begin{align}
\Wind(t) &= \sum_{\tau\le t}\Delta\theta_\tau,\\
\Resist(t) &= \alpha\,\mathbb{E}|\Delta\theta_\tau|/\pi + \beta\,\lambda_t,\quad \lambda_t=\text{returns to prior phases up to }t,\\
\Stress(t) &= \frac{1}{t}\int_0^t w(\tau)\, \Big(D_r(\tau)+\gamma D_\kappa(\tau)\Big)\,d\tau,
\end{align}
with $w(\tau)$ a recency weight and $\gamma$ a curvature weight.
In absence of durations, use a count-based proxy $\rho_{\Jam}=\frac{\#\{\tau: s_\tau=\Jam\}}{t}$ and the simplified $ \Stress \approx \frac{\Wind}{2\pi}\rho_{\Jam}$.

\section{Executable instrumentation (provenance hooks)}
Record in provenance (\texttt{.prov.jsonl}) per step:
(i) phase after transition; (ii) timestamp; (iii) local tensor $\bm{g}_t$ hash or summary; (iv) loop counter $\lambda_t$. Emit end-of-run:
\begin{lstlisting}[language=Python,basicstyle=\ttfamily\small]
{"event":"stress_index",
 "value": S,
 "winding_deg": int(Wind * 180/pi),
 "jam_ratio": rho_JAM,
 "resistance": Resist}
\end{lstlisting}
For time-weighted $\rho_{\Jam}$, log entry/exit timestamps for each phase.

\section{Predictions and falsifiable tests}
\textbf{P1 (monotonicity).} Increasing contradiction density raises $\Wind$ and $\rho_{\Jam}$; hence $\Stress$ rises monotonically.\newline
\textbf{P2 (collapse indicator).} Beyond a threshold in $\Resist$, orbits degenerate near \Jam; manifold Betti-1 persistence shortens (PH signature).\newline
\textbf{P3 (universality).} The $\T^2$ attractor persists across domains (legal, medical, defense) without retuning operators.\newline
\textbf{P4 (beyond biology).} LEE realizes $\T^k$ and altered-genus surfaces; some may later be measured biologically.

\section{Relation to neural-manifold reports (post hoc)}
The resemblance to low-dimensional neural tori \citep{Nature_2011,Nature_2024} was observed \emph{after} LEE had an explicit toroidal phase space.
We emphasize convergence, not inspiration: LEE's torus follows from discrete implication rotations and contradiction archival; any biological parallel is supportive but not formative.

\section{Applications and implications}
\textbf{Diagnostics.} Stress-aware runs separate resolvable contradictions from degenerative loops.\newline
\textbf{Governance.} Provenance + manifold health provides auditable decision trails.\newline
\textbf{Biomedical hypotheses.} LEE can propose manifolds beyond current cortical constraints for empirical probing.

\section{Limitations}
We presently approximate time weights from state counts when phase durations are absent; future engine versions will log exact durations.
Distortion estimators depend on embedding choice (spectral, isomap, PH); we mitigate via method agreement and reporting.

\section{Conclusion}
LEE's toroidal geometry is a law of its logic, not an import from biology. With \GLMS\ and stress-aware metrics, the engine not only performs inference but also exposes a measurable manifold of reasoning—usable for engineering stability and cross-domain science.

\paragraph{Data and code.}
The engine emits JSON, PROV, SVG and (optionally) StressIndex metrics; figures in the arXiv version will be generated directly from these artifacts.

\bibliographystyle{plainnat}
\bibliography{lee_toroidal}
\end{document}
