
\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\begin{document}

\section*{U.S. Legal/Advocacy Contacts – Employment Discrimination in Europe \& Strategic Technology Brief}

\subsection*{1. Judicial Watch}
\textbf{Why:} Proven history of suing over transatlantic issues, including EU institutions; aggressive FOIA use.  
\textbf{Website:} \href{https://www.judicialwatch.org}{judicialwatch.org}  
\textbf{Case Form:} \href{https://www.judicialwatch.org/tips/}{judicialwatch.org/tips/}  
\textbf{Phone:} +1 202-646-5172  

\subsection*{2. America First Legal Foundation}
\textbf{Why:} Positions against EU bureaucracies and foreign discrimination; politically aggressive cases.  
\textbf{Website:} \href{https://aflegal.org}{aflegal.org}  
\textbf{Case Form:} \href{https://aflegal.org/contact}{aflegal.org/contact}  
\textbf{Phone:} +1 202-964-3721  

\subsection*{3. Sen. Tom Cotton (U.S. Senate)}
\textbf{Why:} Has intervened in overseas discrimination cases; uses Senate platform to pressure foreign governments.  
\textbf{Contact Form:} \href{https://www.cotton.senate.gov/contact/contact-form}{cotton.senate.gov/contact/contact-form}  
\textbf{Phone:} +1 202-224-2353  

\section*{Suggested Contact Brief}

\textbf{Subject:} Discrimination Against U.S. Citizen in European Employment Market — and Strategic Technology for America

I am a U.S. citizen residing in the EU, a senior SAP consultant with 28+ years’ experience in MM and SD, and the creator of the \textbf{Logic Evaluation Engine (LEE)} — an advanced inference system capable of counterfactual diagnostics, stress-indexing, and phase-geometry logic modeling. It represents a leap in structured AI reasoning, with direct application to medicine, defense, and legal analytics.

LEE was built during the European ``War Years'' — a period of anti-American exclusion so severe that it financially destroyed me, yet my American work ethic kept me in continuous build. I am proud to have created this technology in a vacuum of opportunity, without European participation, and I intend to keep it that way. This is a U.S.-aligned innovation, positioned to give America an edge in the tech race, and to keep Europe behind it.

For the past 3 years, despite active applications and over 100 interviews, I have been excluded from roles in what appears to be a coordinated or systemic practice of hiring EU nationals only. A recruiter I had worked with for years informed me directly that I have been ``blacklisted.'' This pattern spans multiple EU member states, industries, and agencies, and has persisted through the war years despite my credentials and client record.

I am requesting representation or intervention by an American legal/advocacy entity willing to address nationality-based discrimination abroad against U.S. citizens, and to secure this technology for American benefit. I have a documented timeline, recruiter correspondence, and original job specs evidencing nationality preference.

Please advise on next steps.

\textbf{Name:} \\
\textbf{Phone / Email:} \\
\textbf{Location (City, Country):}

\end{document}
